Structured Versus Decorated Cospans

### 2020-12-02 cospans.tex
small changes in Section 5.4

Chemists sometimes use Petri nets where the places represent `chemical species'---types of molecules, ions and the like---and the transitions represent chemical reactions.   In these Petri nets, which we shall call `Petri nets with rates', the transitions are labelled with positive real numbers called `rate constants'.  This allows one to systematically construct a dynamical system from any Petri net with rates, and mathematical chemists have proved deep theorems relating the topology of Petri nets with rates to the qualitative behavior of their dynamical systems.    Pollard and the first author showed how to construct an `open' dynamical system from an `open' Petri net with rates---meaning one where chemicals can flow in and out at certain places \cite{BP}.  We used the formalism of decorated cospans to do this, and to define a functor from a category with open Petri nets with rates as morphisms to one with open dynamical systems as morphisms.  Later, the first two authors used structured cospans to promote the first of these categories to a double category \cite[Sec.\ 6.16]{BC}. Here we show the second, the category with open dynamical systems as morphism, \emph{cannot} be made into a double category using structured cospans---but we can do it using decorated cospans.

>>>>

Chemists sometimes use Petri nets where the places represent `chemical species'---types of molecules, ions and the like---and the transitions represent chemical reactions.   In these Petri nets, which we shall call `Petri nets with rates', the transitions are labelled with positive real numbers called `rate constants'.  This allows one to systematically construct a dynamical system from any Petri net with rates, and mathematical chemists have proved deep theorems relating the topology of Petri nets with rates to the qualitative behavior of their dynamical systems.    Pollard and the first author showed how to construct an `open' dynamical system from an `open' Petri net with rates, meaning one where chemicals can flow in and out at certain places \cite{BP}.  They used the formalism of decorated cospans to do this, and to define a functor from a category with open Petri nets with rates as morphisms to one with `open dynamical systems' as morphisms.  Later, structured cospans were used to promote the first of these categories to a double category \cite[Sec.\ 6.16]{BC}. Here we show the second, the category with open dynamical systems as morphisms, \emph{cannot} be made into a double category using structured cospans---but we can do it using decorated cospans!
